Formal Verification of Real-Time Data Processing of the LHC Beam Loss Monitoring System: A Case Study

### 2011-03-31 fmics/paper.tex
more dropbox changes

We described a case study in which the HOL4 theorem prover is used to formally verify properties of the SRS structure within the LHC machine protection system.
In this case study, we built a parameterized model of the SRS structure and showed that its behavior is correct with respect to its specification.
It is likely that the configuration of the slices and shift registers will need to change in the future as the understanding of the LHC and its weaknesses increases to accommodate more targeted protections
Thus, the parameterization in our model was crucial to make it applicable to the future upgrades.

@@

The use of mechanized theorem proving to verify the correctness of the SRS behavior complements an intensive verification effort based on simulation performed at CERN.
This effort targeted the validity of the SRS as an accurate and fast enough method by analyzing its behavior (a) in the boundaries of the threshold limits and (b) in expected types of beam losses, e.g., fast and steady losses.
Its results showed that the current implementation of the SRS, given in Figure~\ref{fig:srs}, satisfies the specification, that is the running sums have a maximum 20\% relative error.
On the contrary, the results in this paper apply to all possible input steams, not just a limited number of input stream scenarios performed at CERN.

But, the error bound given in Theorem~\ref{max_error} does not satisfy the maximum 20\% relative error bound.
In order to make this bound tighter, we need to provide more characterization of the input as constraints on $D$.
Due to the nature of the problem, such a characterization of input is not trivial.
Therefore, the SRS structure in LHC is an interesting example in which formal verification does not completely replace conventional verification approaches such as simulation.

HOL4 has highly developed tools  for reasoning about bitstrings (words, bytes etc).
One possible future research direction os to refine the natural number model to a more hardware-realistic one and even verifying a model extracted from the HDL used to synthesize the FPGA.

The general method (I think Mike mentioned) of subjecting a formal
model to "obviousness" tests, collecting small theorems about it along
the way, seems to be a good one.

%Future direction, go with lower level implementation rather than finite words (closer  to vhdl coed) and improve the bound

>>>>

We described a case study in which the HOL4 theorem prover is used to formally verify properties of the SRS structure within the LHC machine protection system.
In this case study, we built a parameterized model of the SRS structure and showed that its behavior is correct with respect to its specification.
It is likely that the configuration of the slices and shift registers will need to change in the future as the understanding of the LHC and its weaknesses increases to accommodate more targeted protections.
Thus, the parameterization in our model was crucial to make it applicable to the future upgrades.

@@

The use of mechanized theorem proving to verify the correctness of the SRS behavior complements an intensive verification effort based on simulation performed at CERN.
This effort targeted the validity of the SRS as an accurate and fast enough method by analyzing its behavior (a) in the boundaries of the threshold limits and (b) in expected types of beam losses, e.g., fast and steady losses.
Its results showed that the current implementation of the SRS, given in Figure~\ref{fig:srs}, satisfies the specification, that is the running sums have a maximum 20\% relative error ($|\mathsf{RS}\;D\;n\;t - \mathsf{true\_sum}\;D\;n\;t| < 20\%(\mathsf{true\_sum}))$.

On the contrary, the results in this paper apply to a larger set of input streams, not just a limited number of sample inputs considered at CERN.
However, the error bound proved in Theorem~\ref{max_error} is not tight enough to satisfy the maximum 20\% relative error.
One of the potential reasons for this problem is the lack of knowledge to characterize the input stream to the SRS.
Due to the nature of this problem, such a characterization of input is not trivial.
While testing and simulation of the SRS on a limited set of input streams offer a level of confidence that the behavior of SRS satisfies this particular specification, constructing a formal proof for all possible behaviors requires a better understanding of the characteristics of the input stream.
One future research direction is to investigate whether constructing such a formal proof is feasible for the SRS component of BLMS.

Another future research direction is to refine our formal model to a lower level of abstraction and closer to the hardware model (e.g., representing the count values by words instead of natural number).
We are also considering verifying a model extracted from the Hardware Description Language (HDL) used to synthesize the BLETC FPGA.
